BOLD Parameter Estimation Using Particle Filters

### 2010-07-19 doc/thesis/results.tex
first draft done

\section{Multi-voxel Simulation}
To test the usefullness of the particle filter for Second I used a modified version of the FSL tool 
POSSUM to generate an entire FMRI image from a parameter map. The parameter map was generated
by taking an existing activation map and assigning discrete parameter sets to each region.
The result was a four dimensional (length x width
x height x parameter) image with spatially varying parameters. Possum was then modified
to take a parameter map and generate activation levels depending on the parameters at that
point. The patch for POSSUM will be made available. As an unfortunate side effect of 
not using Possums' original activation scale, I manually added 750 to the total level of
simulated Possum images. This is because the BOLD \% difference levels were in the range
of 50 - 100\% from the base, about 5 times as large as they should have been. Ultimately
this should not have an effect on the parameters (other than perhaps $\epsilon$ and $V_0$). 
Also worth noting is that the noise level was set to an SNR of 20, due to changes in the 
program the true signal-to-noise ratio was much lower, as soon in the fit in \autoref{fig:simslicefit1}.

\begin{figure}
%includegraphics[]{images/simslicefit1}
\caption{A single time-series fit in the simulated slice.}
\labe{fig:simslicefit1}
\end{figure}

@@

\begin{figure}
\includegraphics[]{images/simslice_hm}
\caption{Normalized Square Root of the Mean Square Residual. Darker indicates lower error.}
\label{fig:simslice_hm}
\end{figure}

\begin{figure}
\includegraphics[]{images/simslice_hm85}
\caption{Normalized Square Root of the Mean Square Residual, with upper threshold at .85. 
Darker indicates lower error.}
\label{fig:simslice_hm85}
\end{figure}

\begin{figure}
\includegraphics[]{images/parameters_sec1}
\caption{Histogram of estimated parameters in section with true parameters of $\{ \}$
Limited to areas with residual less than $.85$.}
\label{fig:parameters_sec1}
\end{figure}

\begin{figure}
\includegraphics[]{images/parameters_sec2}
\caption{Histogram of estimated parameters in section with true parameters of $\{ \}$
Limited to areas with residual less than $.85$.}
\label{fig:parameters_sec2}
\end{figure}

\begin{figure}
\includegraphics[]{images/parameters_sec3}
\caption{Histogram of estimated parameters in section with true parameters of $\{ \}$. 
Limited to areas with residual less than $.85$.}
\label{fig:parameters_sec3}
\end{figure}

>>>>

\section{Multi-voxel Simulation}
To test the usefullness of the particle filter for Second I used a modified version of the FSL tool 
POSSUM to generate an entire FMRI image from a parameter map. The parameter map was generated
by taking an existing activation map and assigning discrete parameter sets to each region.
The result was a four dimensional (length x width
x height x parameter) image with spatially varying parameters. Possum was then modified
to take a parameter map and generate activation levels depending on the parameters at that
point. The patch for POSSUM will be made available. As an unfortunate side effect of 
not using Possums' original activation scale, I manually added 750 to the total level of
simulated Possum images. This is because the BOLD \% difference levels were in the range
of 50 - 100\% from the base, about 5 times as large as they should have been. Ultimately
this should not have an effect on the parameters (other than perhaps $\epsilon$ and $V_0$). 
Also worth noting is that the noise level was set to an SNR of 20, due to changes in the 
program the true signal-to-noise ratio was much lower, as seen in the fit in \autoref{fig:simslicefit1}.

\begin{figure}[H]
\centering
\includegraphics[width=15cm]{images/simslice_noise}
\caption{A single time-series fit in the simulated slice.}
\label{fig:simslicefit1}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[width=10cm]{images/simslice_mask}
\caption{Normalized Square Root of the Mean Square Residual. Darker indicates lower error.}
\label{fig:simslice_mask}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[width=10cm]{images/simslice_hm}
\caption{Normalized Square Root of the Mean Square Residual. Darker indicates lower error.}
\label{fig:simslice_hm}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[width=10cm]{images/simslice_hm85}
\caption{Normalized Square Root of the Mean Square Residual, with upper threshold at .85. 
Darker indicates lower error.}
\label{fig:simslice_hm85}
\end{figure}

The regions are numbered according to \autoref{fig:simslice_mask}; the parameters
for each region may be found in \autoref{tab:simsliceparams}.

\begin{table}[t]
\centering
\begin{tabular}{|c |c | c | c | c | c | c | c |}
\hline 
Region & $\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  \\
\hline 
1 & 1.454& .321& .369& .036& .994& 2.774& 1.348\\
2 &1.151& .353& .380& .026& 1.98& 2.333& 1.645 \\
3 &1.951& .317& .348& .027& 1.657& 3.719& .757 \\
4 &1.203 & .310& .326& .036& 2.168& 2.272& .086\\
\hline
\end{tabular}
\caption{Actual parameters for each regions in the simulated slice.}
\label{tab:simsliceparams} 
\end{table}
%last (bottom right), 1.203, .310, .326, .036, 2.168, 2.272, .086, 3.4, 1

Note that region 4 has a very low $\epsilon$. For this reason, the only areas
with significantly decent estimates of the BOLD time series are 1,2 and 3. With
such a low $\epsilon$ region 4 was below the noise threshold. Notice that the
regions 1 2 an 3 clearly stick out in the $\sqrt{MSR}$ map, indicating that the
particle filter was succesfull in matching those regions.

\begin{figure} %bottom left
\centering
\includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=15cm]{images/slicesim_hist1}
\caption{Histogram of estimated parameters in section 1 with $\sqrt{MSR}$ less than $.85$.}
\label{fig:slicesim_hist1}
\end{figure}

\begin{figure} %top left
\centering
\includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=15cm]{images/slicesim_hist2}
\caption{Histogram of estimated parameters in section 2 with $\sqrt{MSR}$ less than $.85$.}
\label{fig:slicesim_hist2}
\end{figure}

\begin{figure} %top right
\centering
\includegraphics[clip=true,trim=2.5cm 2cm 2cm 1cm,width=15cm]{images/slicesim_hist3}
\caption{Histogram of estimated parameters in section 3 with  $\sqrt{MSR}$ less than $.85$.}
\label{fig:slicesim_hist3}
\end{figure}

It is again clear that a single point estimate of the parameters is not feasible,
but rather a distribution of parameters is the best any modeling algorithm
is going to be capable of for this version of the BOLD signal. In spite of
this, and relatively heavy noise, the fact that is that the fits were
good. Clearly, from the normalized $\sqrt{MSR}$ it is possible to identify
 \"active" regions; or regions that respond the way the BOLD model expects.
Although some of the fits in the  \"active" regions did not have good regression,
this is not necessarily the work of false negatives. Instead, because possum
simulates different tissues, it is more likeley that those regions non
grey-matter areas. Essentially these areas should not have reacted very strongly
because in the real world such regions wouldn't. By the fact that it 
was relatively easy to identify the regions of activation from the simulated 
results, something similar should happen with real FMRI data, which is discussed
in the next session. which
